---
title: "querying-measurements"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{querying-measurements}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(openaq)
```

## Introduction

This vignette demonstrates how to use the openaq R package to query air quality
measurements from the OpenAQ API. We will walk through the process of retrieving
a list of locations, inspecting their sensors and parameters, and finally,
fetching and plotting the actual measurements for a specific sensor.


### Listing Locations

The `list_locations()` function allows us to retrieve information about
monitoring locations. We can filter locations based on parameters and countries.
Here, we are searching for locations that measure PM<sub>2.5</sub>
(parameter ID 2) in Madagascar (country ID 182). By default resource functions
return a data frame but we set `as_data_frame = FALSE` to receive the data as a
list, which we will use for easier inspection in the following steps.

```{r eval=FALSE}
locations <- list_locations(
  parameters_id = 2,
  countries_id = 182,
  as_data_frame = FALSE
)
```


The locations object is a list, where each element represents a monitoring
location. Let's examine the first location in the list and explore its sensors.

```{r eval=FALSE}

location <- locations[[1]]
sensors <- location$sensors
sensors
# > [[1]]
# > [[1]]$id
# > [1] 225221
# >
# > [[1]]$name
# > [1] "pm25 µg/m³"
# >
# > [[1]]$parameter
# > [[1]]$parameter$id
# > [1] 2
# >
# > [[1]]$parameter$name
# > [1] "pm25"
# >
# > [[1]]$parameter$units
# > [1] "µg/m³"
# >
# > [[1]]$parameter$displayName
# > [1] "PM2.5"
```


Each location contains information about its sensors. Each sensor has an id,
name, and details about the measured parameter, including its name, units, and
ID.


Each location also provides the timezone of where it is local, which will be
helpful when querying measurements, so we will store this for later.

```{r eval=FALSE}
tz <- location$timezone
tz
# > [1] "Indian/Antananarivo"
```


To fetch measurements, we need the sensor ID. This location only measures
PM<sub>2.5</sub> and in turn only has one sensor so we can extract it as
follows:

```{r eval=FALSE}

sensors_id <- sensors[[1]]$id
sensors_id
# > [1] 225221
```

## Fetching Measurements

Now we can use the `list_sensor_measurements()` function to retrieve the
measurements for the sensor. We need to provide the `sensors_id`, 
`datetime_from`, and `datetime_to` arguments. We will query for all measurements
during the month of January 2025.It is important to use the correct timezone for
your dates, we use the timezone for location that we stored in the `tz` variable
above. The limit argument controls the maximum number of measurements returned
in a single page of results. Because this is hourly data, and we are querying 31
days of measurements, we will be able to view all the results in one page (24
hourly measurements * 31 days = 744 measurements).

```{r eval=FALSE}
measurements <- list_sensor_measurements(
  sensors_id,
  datetime_from = as.POSIXct("2025-01-01", tz = tz),
  datetime_to = as.POSIXct("2025-01-31", tz = tz),
  limit = 1000
)
```

Finally, we can visualize the measurements using the `plot()` function.

```{r eval=FALSE}
plot(measurements)
```
